Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-11-01 doc/paper-lcr.tex
abstract sharpened up

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight.
Such massive structures contribute an ``external convergence'' $\kappa$ that
in some cases dominates the uncertainty in the inferred distances.  We
characterise this uncertainty by marginalising over a probability density
function for $\kappa$: here, we investigate the use of a simple halo model to
estimate $\pr(\kappa|\data)$ given noisy estimates of the photometric redshift
and stellar mass of galaxies observed in optical imaging surveys of the field
in question. We use mock catalogs from the Millennium Simulation, and compare
our reconstructed $\pr(\kappa|\data)$ to ray-traced $\kappa$ truth values: 
taking into account realistic uncertainties on redshift and stellar masses, we
typically find that the reconstruction yields an improvement in precision of
40\% over the global $\pr(\kappa)$. For any given line of sight the
convergence estimate can be made before the investment of expensive follow-up
observations: we find that selecting systems by the precision with which
$\kappa$ can be estimated leads to a low density sample that would give
distance estimates a factor of 2 more precise, with biases as low as 0.5\%.
Preferentially selecting lines of sight with high external shear (as might
happen in an assembled sample of quadruply-imaged time delay lenses) has the
opposite effect, with distances biased by around 1\%. We suggest areas for the
improvement of this general analysis framework that should allow yet more
accurate cosmological inferences to be made.

>>>>

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight.
Such massive structures contribute an ``external convergence'' $\kappa$ that
in some cases dominates the uncertainty in the inferred distances.  We
characterise this uncertainty by marginalising over a probability density
function for $\kappa$: here, we investigate the use of a simple halo model to
estimate $\pr(\kappa|\data)$ given noisy estimates of the photometric redshift
and stellar mass of galaxies observed in optical imaging surveys of the field
in question. We use mock catalogs from the Millennium Simulation, and compare
our reconstructed $\pr(\kappa|\data)$ to ray-traced $\kappa$ truth values: 
taking into account realistic uncertainties on redshift and stellar masses, we
typically find that the reconstruction yields an improvement in precision of
40\% over the global $\pr(\kappa)$. For any given line of sight the
convergence estimate can be made before the investment of expensive follow-up
observations: we find that selecting systems by the precision with which
$\kappa$ can be estimated leads to a low convergence  sample that would give
time delay distance estimates a factor of 2 more precise than a random
selection,  with biases as low as 0.3\%. Preferentially selecting lines of
sight with high external shear (as might happen in an assembled sample of
quadruply-imaged time delay lenses) has the opposite effect, with distances
biased by around 1\%. We suggest areas for the improvement of this general
analysis framework that should allow yet more accurate cosmological inferences
to be made.
